The Early Eocene Climate Optimum reconstructed with Bayesian multi-proxy integration

### 2022-07-22 manuscript/manuscript.Rmd
add notes to ms
describe mangroves as a proxy

@@

@@

## Fig 1 - Model structure and proof of concept with modern data
## Fig 2 - Eocene example
## Fig 3 - Showing influence of changing model assumptions

>>>>

In order to properly compare model output to proxy data, latitudinal temperature patterns need to be reconstructed from noisy data (impossible to properly compare model gradients with scattered proxy points "by eye")

@@

We chose a range of proxies for marine, shallow water temperatures
proxies: d18O, D47, Mg/Ca, TEX86
mangroves, corals, sharks?

## early Eocene SST proxies ##

**d18O.**

**d47.**

**Mg/Ca.**

**TEX86.**

**Mangroves.** Recent mangroves predominantly occur in the tropics - subtropics. Although many factors besides temperature determine the extent of mangrove distributions, empirical temperature limits have been established for the genera *Avicennia* and *Rhizophora*, with lower mean annual SST limits of 15.6 degC and 20.8 degC, respectively [@Quisthoudt2012]. Both *Avicennia* and members of the Rhizophoraceae family were widespread in the early Eocene, but only *Avicennia* occurred in polar latitudes [@Suan2017; @Popescu2021]. Assuming that Eocene members of these mangrove taxa observed similar climatic requirements as their modern relatives, the presence and absence of *Avicennia* and Rhizophoraceae pollen can be used as a temperature proxy. To translate mangrove occurrences from the compilation by @Popescu2021 to a quantitative temperature proxy, we allocate locations at which *Avicennia* pollen, but no Rhizophoraceae pollen were found, a mean temperature between the lower *Avicennia* and Rhizophoraceae limit (18.2 degC), with a standard deviation of 1.33 degC, thus placing 95 % of the probability density within that interval (see Fig. 1x). For locations at which both *Avicennia* and Rhizophoraceae pollen were found, the range of plausible temperatures is very wide and was not used in the analysis, as both mangrove taxa occur in temperatures > 32 deC in the modern.

**Corals.**

[Fig 1: In this part we can show a figure with the proxy data and the distributions of temperature proxies we specify, like mangroves]

## A hierarchical model of latitudinal temperature gradients ##

[Fig 2: In this part we can show a figure with the model structure and the priors we put on the parameters. Also, try it on the modern data? Alternatively, this figure goes to the SI]

@@

## Fig 3 - Eocene gradient
## Fig 4 - compare Eocene gradient with previous reconstructions / ES models
